---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

### Loading

Let's check if we have a "data" folder in our analysis template. Then we'll upload and unzip the data there.

```{r cache=TRUE}

# Does this directory exist. If not, I'll do all the work for you.
if(!file.exists("./data")) { dir.create("./data") }

# The zip file.
fileURL <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"

# Downloading the zip and placing it in the data folder.
download.file(fileURL,"./data/repdata-data-activity.zip", method="curl")

# Unzip and TADA!
unzip("./data/repdata-data-activity.zip", exdir="data")

# Let's get rid of that ugly zip file because I like to keep things clean.
file.remove("./data/repdata-data-activity.zip")

# Last stop: data-ville
data <- read.csv("./data/activity.csv")

```

### Preprocessing

Let's take a peek under the data's hood and see what we're dealing with.

```{r}

str(data)

```

That "date" variable should be in the Date class to make life easier for everyone. I prefer dplyr's mutate function because it makes me feel like I'm creating a new X-men character who has the power to report dates accurately.

```{r message=FALSE}

library(dplyr)

data <- data %>% mutate(date = as.Date(date))


```


## What is mean total number of steps taken per day?

Let's create a plot of the overall steps taken per day and then find our mean and median.

```{r warning=FALSE}
library(ggplot2)

ggplot(data = data) + geom_bar(aes(date, steps), stat="identity")

```

## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
